Forecasting the 2024 U.S. Presidential Election Using a Model-Based Approach

### 2024-11-03 paper/paper.qmd
update paper

---
title: "Forecasting the 2024 U.S. Presidential Election Using a Model-Based Approach"
subtitle: "Harris expects to win on 49%, leads Trump by about 2% in 2024 US election"
author: 
  - Diana Shen
  - Jinyan Wei
  - Jerry Yu
thanks: "Code and data are available at: [https://github.com/DianaShen1224/Forecast-2024-US-election](https://github.com/DianaShen1224/Forecast-2024-US-election)."
date: today
date-format: long
abstract: "This paper predicts the outcome of the 2024 U.S. presidential election using a statistical model based on aggregated polling data for Kamala Harris and Donald Trump. We employ multi-level regression with post-stratification (MRP) using demographic predictors to estimate voter support. The analysis addresses polling biases and proposes an idealized survey methodology with a $100,000 budget. Our results offer insights into voter behavior and suggest improvements for future election forecasting models."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

@@

library(tidyverse)
library(arrow)
library(dplyr)
library(ggplot2)
library(modelsummary)
library(RColorBrewer)
library(gt)
library(future)
```

@@

Research by Gelman and King (1993) underscores that while polls can
exhibit variability over a campaign period, they can provide reliable
predictions when adjusted for temporal changes and fundamental factors.
Additionally, Erikson and Wlezien (2008) highlight the importance of
timing in polling, showing that data closer to election day tends to
stabilize and therefore becomes more predictive. Our research seeks to
fill a gap in existing literature that often fails to address the
complexities of polling data, thereby enhancing our understanding of
voter behavior within the electoral context.

@@

The structure of this paper is organized as follows: @third-data
provides details on the data sources and variables used in our analysis.
@forth-model explains the modeling approach, including the assumptions
and specifications of our linear regression framework. In
@fifth-results, we present our findings, emphasizing the key predictors
of Harris's support. Finally, @sixth-discussion explores the
implications of our results and suggests potential directions for future
research. @appendix-a provide external data detail, @appendix-b provide
model detail, @appendix-c provide a exploration for pollster
methodology, @appendix-d provide a idealized methodology.

@@

We conduct our polling data analysis using the R programming language
[@citeR]. Our dataset, obtained from FiveThirtyEight
[@fivethirtyeight2024],based on polling as of 27 October 2024, provides
a detailed overview of public opinion in the lead-up to the election.
Adhering to the guidelines presented in @tellingstories, we explore
various factors that influence voter support percentages, including the
timing of the polls, the traits of polling organizations, and regional
differences.

In this study, we utilized several R packages to enhance our data
manipulation, modeling, and visualization capabilities. The tidyverse
package offered a comprehensive set of tools for data wrangling and
analysis, improving workflow efficiency [@thereferencecanbewhatever].
The here package aided in managing file paths, allowing for easy access
to our data files [@citehere]. We relied on janitor to perform data
cleaning, as it provides functionalities to identify and rectify quality
issues within the dataset [@citejanitor]. For handling date-related
operations, the lubridate package proved invaluable, simplifying the
manipulation of time variables [@citelubridate]. Lastly, arrow supported
efficient data input and output in a performance-oriented format,
essential for managing larger datasets [@citearrow]. Our coding
practices and file organization were informed by the structure outlined
in @tellingstories.

@@

@@

```{r}
#| label: fig-percentage
#| fig-cap: Distribution of support percentage of Kamala Harris
#| echo: false
#| warning: false
#| message: false
ggplot(harris_data, aes(x = pct)) +
  geom_histogram(binwidth = 1, fill = "gold", color = "grey50") +
  labs(
    x = "Support Percentage",
    y = "Count"
  ) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
```

@@

The end date is the final day of data collection for a poll, indicating
when the survey period concluded. This date provides crucial context for
the poll results, as public opinion may change over time in response to
events, campaign actions, or other influencing factors.

## Recency {#sec-recency}

Recency is a newly created predictor variable in our dataset, specifically
for use in our model. It is calculated as follows Recency = Election Date - End Date,
where Election Date is set to November 5th. This variable quantifies the time
interval between the poll's end date and the election date as a measure of the poll's
relevance and potential impact on voter behavior. In general, polls conducted closer
to the election date provide a more accurate picture of public opinion. The original
dataset includes only dates, and by mutating this variable, we can more intuitively
interpret the number of days before the election that each data point was collected.
Converting this to discrete data enhances its usefulness in constructing predictive models.

## National Poll {#sec-national-poll}

National polls is another newly created variable in our dataset, represented by a binary
indicator (1 for national polls, 0 otherwise). This distinction is important because
not all polls are conducted nationally; some are state-specific and may reflect the
region's preference for a particular presidential candidate. Ignoring this distinction
may introduce significant bias into our model. By labeling each poll as national or
non-national, we can give higher weight to national polls, which are more representative
of overall public opinion, and thus improve the accuracy of our predictive model.

## Population {#population}

The “population” variable in the dataset represents the status of each
entry. it is the abbreviated description of the respondent group, categorizing
participants as “lv” (likely voters) or “rv” (registered voters). They represent
different levels of participation and potential voting behavior. Likely voters
are individuals who are not only registered but are statistically more likely
to vote based on historical data or survey responses. In contrast, registered
voters include all individuals who are eligible to vote, regardless of their
likelihood of participation.

@@

```{r}
#| label: fig-state
#| fig-cap: Count of polls by state
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-pollster
#| fig-cap: Frequency of Polls by Pollster"
#| echo: false
#| warning: false
#| message: false

@@

The linear regression models constructed for predicting support for
Kamala Harris and Donald Trump in the 2024 U.S. Presidential Election
utilized both unweighted and weighted configurations.The model summary of support for Harris is shown in @tbl-summary1, support for Trump is shown in @tbl-summary2. As shown in
@tbl-summary1 and @tbl-summary2, the intercepts for Harris and Trump,
estimated at 44.70 and 59.28 respectively in the unweighted models, while in weighted models, it's estimated at 45.2 and 54.8,
represent their baseline support levels when all predictors are at
reference values. The high ( R\^2 ) values of 0.824 for Harris’s
unweighted model and 0.892 for her weighted model, along with similar
values of 0.849 and 0.916 for Trump, indicate that a significant portion
of the variation in support can be explained by the model's predictors.
The relatively low Root Mean Square Error (RMSE) values further suggest
that the models provide reasonably accurate predictions of support
levels.

@@

@fig-support-trend illustrate the predicted support for Kamala Harris
and Donald Trump from August 1, 2024, to October 27, 2024, using both
unweighted and weighted linear regression models. In the
@fig-support-trend-1 unweighted model, we see that Trump's predicted
support generally hovers between 40% and 50%, while Harris's support
ranges from 45% to 50%. Both candidates' support levels appear
relatively stable across the polling period, with minor fluctuations.
This model, which treats all polls equally, suggests a consistent lead for Harris, albeit within a fairly close range.

The weighted model @fig-support-trend-2 refines these predictions by
applying weights based on recency and sample size, which reduces the
influence of older or smaller polls. This approach results in smoother
trend lines for both candidates, with less apparent variability compared
to the unweighted model @fig-support-trend-1. Trump's support remains
around the 40% to 50% range, while Harris’s support stays in about 45% to 52%  range. The weighting emphasizes more recent and larger polls, likely
providing a more accurate reflection of current public opinion. Overall, both models consistently indicate a lead for Harris, with the weighted model slightly stabilizing the predictions. The average support for Harris and Trump are both consist for two models, which predicts that Harris will win on about 49%, leads Trump by about 2%.

@@

set.seed(233)
plan(multisession)
####### Prediction for Harris ######
new_data_harris <- data.frame(
  end_date = seq(
    min(harris_data$end_date),
    max(harris_data$end_date),
    length.out = 700
  ),
  state = factor(sample( c("National","Arizona", "Georgia", "Michigan", "Nevada", "North Carolina", "Pennsylvania", "Wisconsin"), size = 700, replace = TRUE)), 
  pollster = factor(sample(levels(harris_data$pollster), size = 700, replace = TRUE), 
                    levels = levels(harris_data$pollster)), 
  population = factor(sample(levels(harris_data$population), size = 700, replace = TRUE), 
                    levels = levels(harris_data$population)),
  sample_size = sample(1:2300, size = 700, replace = TRUE)
)|>
  mutate(
    national_poll = if_else(state == "National", 1, 0),
    recency = as.numeric(difftime(as.Date("2024-11-05"), end_date, units = "days")),
    sample_size_weight = pmin(sample_size / 2300, 1),
    recency_weight = exp(-recency * 0.1),
    combined_weight = recency_weight * sample_size_weight,
    candidate="Kamala Harris"
  )

####### Prediction for Trump ######
new_data_trump <- data.frame(
  end_date = seq(
    min(trump_data$end_date),
    max(trump_data$end_date),
    length.out = 700
  ),
  pollster = factor(sample(levels(trump_data$pollster), size = 700, replace = TRUE), 
                    levels = levels(trump_data$pollster)),
 population= factor(sample(levels(trump_data$population), size = 700, replace = TRUE), 
                    levels = levels(trump_data$population)),
   state = factor(sample( c("National","Arizona", "Georgia", "Michigan", "Nevada", "North Carolina", "Pennsylvania", "Wisconsin"), size = 700, replace = TRUE), 
                    levels = levels(trump_data$state)),  
  sample_size =sample(1:2300, size = 700, replace = TRUE)
)|>
  mutate(
   national_poll = if_else(state == "National", 1, 0),
      recency = as.numeric(difftime(as.Date("2024-11-05"), end_date, units = "days")),
    sample_size_weight = pmin(sample_size / 2300, 1),
    recency_weight = exp(-recency * 0.1),  
    combined_weight = recency_weight * sample_size_weight,
    candidate="Donald Trump"
  )
```

@@

# Filter combined_predictions to include only data from the recent date cutoff onward
combined_predictions<- bind_rows(new_data_harris, new_data_trump)
combined_predictions_filtered<-combined_predictions|>filter(end_date>=recent_date_cutoff)
# Determine the maximum end date for setting the x-axis limit
max_end_date <- as.Date("2024-10-26")
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| fig-cap: "Harris Kamala leads Donald Trump an average 5% in Predicted Support in the 2024 US Election (From August 1, 2024, to October 27, 2024)"
#| fig-subcap: ["Unweighted","Weighted"]
#| layout-ncol: 2
#| label: fig-support-trend
# Plotting with the filtered data
plan(multisession)
##unweighted
ggplot(combined_data_total,aes(x = end_date,y=pct, colour = candidate)) +
  geom_point(alpha = 0.1)+ 
  geom_smooth(data = combined_predictions_filtered,aes(x=end_date,y = predicted_unweighted), method = "lm",size = 1)+ 
  geom_line(data = combined_predictions_filtered,aes(x=end_date,y = predicted_unweighted),size = 0.2)+
  labs(
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate",caption = 
                   "Linear Regression Model: Candidate Support Percentage = national_poll + pollster + population + state") +
  scale_color_manual(values = c("Kamala Harris" = "blue", "Donald Trump" = "red")) +
  theme_classic() +
  theme(legend.position = "right",
        axis.text.x = element_text(angle = 45, hjust = 1),strip.background = element_rect(colour = "black", fill = "white")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date)) 
##weighted
ggplot(combined_data_total,aes(x = end_date,y=pct, colour = candidate,size = combined_weight)) +
  geom_point(alpha = 0.1)+ 
  geom_smooth(data = combined_predictions_filtered,aes(x=end_date,y = predicted_unweighted), method = "lm",size = 1)+ 
  geom_line(data = combined_predictions_filtered,aes(x=end_date,y = predicted_weighted), size = 0.2)+
  labs(x = "End Date of the Poll",
       y = "Predicted Support (%)",
       color = "Candidate",
       size="Weight",
       caption = "Weights: recency_weight * sample_size_weight \nLinear Regression Model: Candidate Support Percentage = national_poll + pollster + population + state") +
  scale_color_manual(values = c("Kamala Harris" = "blue", "Donald Trump" = "red"))  +
  theme_classic() +
  theme(legend.position = "right",
        axis.text.x = element_text(angle = 45, hjust = 1),strip.background = element_rect(colour = "black", fill = "white")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date))
```

@@

@fig-support-state compare the predicted support for Kamala Harris and
Donald Trump in key battleground states (Arizona, Georgia, Michigan,
Nevada, North Carolina, Pennsylvania, and Wisconsin) for the 2024 U.S.
election. @fig-support-state-1 uses an unweighted multiple linear
regression model, while @fig-support-state-2 applies a weighted model,
giving greater influence to polls with higher weights based on recency
and sample size. In both figures, Trump’s support, represented by red
lines, and Harris’s support, shown in blue, show close competition
across most states, with slight variations in trends across the polling
period from August to October 2024.

@@

The @fig-support-state-2 stabilizes these trends by emphasizing recent,
larger polls. Harris's support becomes more consistent, particularly in
Pennsylvania, Michigan, and Wiscosin, where she shows a clearer
lead. This weighted approach provides a more refined snapshot,
highlighting Harris’s likely advantage in these key states.

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| fig-subcap: ["Unweighted","Weighted"]
#| layout-ncol: 2
#| fig-cap: "Harris is likely to lead Trump in Pennsylvania, North Carolina, and Wisconsin by Model Prediction"
#| label: fig-support-state
## unweighted
plan(multisession)
selected_states <- c("Arizona", "Georgia", "Michigan", "Nevada", "North Carolina", "Pennsylvania", "Wisconsin")
combined_data_total<-combined_data_total|>filter(state %in% selected_states)
combined_predictions_filtered<-combined_predictions_filtered|>filter(state %in% selected_states)

ggplot(combined_data_total,aes(x = end_date,y=pct, colour = candidate)) +
  geom_point(alpha = 0.1)+ 
  geom_line(data = combined_predictions_filtered,aes(x=end_date,y = predicted_unweighted), size = 0.5)+ # Smoothed trend line
  facet_wrap(~ state,ncol = 4) +
  labs(
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate",
       caption = "Candidate Support Percentage = national_poll + pollster + population + state") +
  scale_color_manual(values = c("Kamala Harris" = "blue", "Donald Trump" = "red")) +
  theme_classic() +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 90, hjust = 1),
        legend.box = "vertical",
        strip.background = element_rect(colour = "black", fill = "white")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date))

## weighetd
ggplot(combined_data_total,aes(x = end_date,y=pct, colour = candidate,size = combined_weight)) +
  geom_point(alpha = 0.1)+ 
  geom_line(data = combined_predictions_filtered,aes(x=end_date,y = predicted_weighted), size = 0.5)+ # Smoothed trend line
  facet_wrap(~ state,ncol = 4) +
  labs(
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate",
       size="Weight",
       caption = "Weights: recency_weight * sample_size_weight\nCandidate Support Percentage = national_poll + pollster + population + state") +
  scale_color_manual(values = c("Kamala Harris" = "blue", "Donald Trump" = "red")) +
  theme_classic() +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 90, hjust = 1),
        legend.box = "vertical",
        strip.background = element_rect(colour = "black", fill = "white")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date)) 
```

@@

# Discussion

@@

>>>>

---
title: "Forecasting the 2024 U.S. Presidential Election Using a Model-Based Approach"
subtitle: "Kamala Harris expects to win on 47.5%, leads Donald Trump by about 1% in 2024 US election"
author: 
  - Diana Shen
  - Jinyan Wei
  - Jerry Yu
thanks: "Code and data are available at: [https://github.com/DianaShen1224/Forecast-2024-US-election](https://github.com/DianaShen1224/Forecast-2024-US-election)."
date: today
date-format: long
abstract: "This study models the 2024 U.S. Presidential Election by analyzing polling data with adjustments for factors like recency and sample size, simulating an approach that captures up-to-date voter sentiment. Using recent 4 months aggregated poll data, the model identifies shifts in candidate support across key battleground states, providing a nuanced view of evolving preferences as election day approaches. By emphasizing recent, high-quality polls, the analysis improves predictive reliability, offering insights that could inform campaign strategies and public discourse. These findings highlight the value of weighting and time-sensitive adjustments in forecasting methods, balancing accuracy with responsiveness to the latest trends."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

@@

library(tidyverse)
library(arrow)
library(dplyr)
library(ggplot2)
library(modelsummary)
library(RColorBrewer)
library(future)
library(boot)    # For bootstrapping
library(caret)   # For cross-validation
```

@@

Research by @GelKin93 underscores that while polls can
exhibit variability over a campaign period, they can provide reliable
predictions when adjusted for temporal changes and fundamental factors.
Additionally, @politic highlight the importance of
timing in polling, showing that data closer to election day tends to
stabilize and therefore becomes more predictive. Our research seeks to
fill a gap in existing literature that often fails to address the
complexities of polling data, thereby enhancing our understanding of
voter behavior within the electoral context.

@@

The structure of this paper is organized as follows: @sec-data 
provides details on the data sources and variables used in our analysis.
@sec-model explains the modeling approach, including the assumptions
and specifications of our linear regression framework. In
@sec-result, we present our findings, emphasizing the key predictors
of Harris's support. Finally, @sec-discussion explores the
implications of our results and suggests potential directions for future
research. @sec-a provide external data detail, @sec-b provide
model detail, @sec-c provide a exploration for pollster Siena College's
methodology, and  @sec-d provide a idealized methodology for polling the poll.

@@

We conduct our polling data analysis using the R programming language
[@citeR]. Our dataset, obtained from FiveThirtyEight
[@fivethirtyeight2024],based on polling as of 2 November 2024, provides
a detailed overview of public opinion in the lead-up to the election.
Adhering to the guidelines presented in @tellingstories, we explore
various factors that influence voter support percentages, including the
timing of the polls, the traits of polling organizations, and regional
differences.

In this study, we utilized several R packages to enhance our data
manipulation, modeling, and visualization capabilities. The tidyverse
package offered a comprehensive set of tools for data wrangling and
analysis, improving workflow efficiency [@thereferencecanbewhatever].
The here package aided in managing file paths, allowing for easy access
to our data files [@citehere]. We relied on janitor to perform data
cleaning, as it provides functionalities to identify and rectify quality
issues within the dataset [@citejanitor]. For handling date-related
operations, the lubridate package proved invaluable, simplifying the
manipulation of time variables [@citelubridate]. Arrow supported
efficient data input and output in a performance-oriented format,
essential for managing larger datasets [@citearrow]. We use ggplot2 [@citeggplot2] to visualize the analysis of data. In addition to the core packages, we employed several specialized libraries to facilitate specific analytical tasks. The modelsummary package enabled us to generate clean, interpretable model summaries, streamlining the reporting of regression and other statistical results [@citemodelsummary]. RColorBrewer provided color palettes for visualizations, enhancing clarity and visual appeal in our plots [@citeRColorBrewer]. The future package supported parallel processing, optimizing the runtime of intensive computations, especially beneficial in model training [@citefuture]. Finally,for resampling and bootstrapping techniques, we utilized the boot package, which allowed us to assess model stability and reliability [@citeboot1;@citeboot2]. caret facilitated model training and validation, offering a unified interface for cross-validation and performance metrics [@citecaret]. Our coding practices and file organization were informed by the structure outlined
in @tellingstories.

@@

## Variables

The end date is the final day of data collection for a poll, indicating
when the survey period concluded. This date provides crucial context for
the poll results, as public opinion may change over time in response to
events, campaign actions, or other influencing factors. In our analysis, we only choose the polls with end date after July 21 2024, which Kamala Harris first declared in the election.

### Recency {#sec-recency}
```{r}
#| label: fig-recency
#| fig-cap: Distribution of the recency of the polls after July 21 2024
#| echo: false
#| warning: false
#| message: false
ggplot(harris_data, aes(x = recency)) +
  geom_histogram(binwidth = 1, fill = "gold", color = "grey50") +
  labs(
    x = "Recency (days)",
    y = "Count"
  ) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
```

Recency is a newly created predictor variable in our dataset, specifically
for use in our model. It is calculated as follows Recency = Election Date - End Date,
where Election Date is set to November 5th. This variable quantifies the time
interval between the poll's end date and the election date as a measure of the poll's
relevance and potential impact on voter behavior. In general, polls conducted closer
to the election date provide a more accurate picture of public opinion. The original
dataset includes only dates, and by mutating this variable, we can more intuitively
interpret the number of days before the election that each data point was collected.
Converting this to discrete data enhances its usefulness in constructing predictive models.
@fig-recency reveals a strong focus on recent polling data, with the highest frequency of polls occurring within the first few days and gradually declining as recency increases. Periodic peaks around 30, 60, and 90 days suggest some regularity in polling schedules, though the overall trend emphasizes newer data. This distribution highlights the prioritization of recent polls in the dataset.

@@

```{r}
#| label: fig-percentage
#| fig-cap: Distribution of support percentage of Kamala Harris
#| echo: false
#| warning: false
#| message: false
ggplot(harris_data, aes(x = pct)) +
  geom_histogram(binwidth = 1, fill = "gold", color = "grey50") +
  labs(
    x = "Support Percentage (%)",
    y = "Count"
  ) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
```

@@

```{r}
#| label: tbl-category
#| tbl-cap: Summary Statistics of categorical variable population and national_pull
#| echo: false
#| warning: false
#| message: false
harris_data|>select(population,national_poll)|>
  datasummary_skim(type = "categorical")
```

### National Poll {#sec-national-poll}

National polls is another newly created variable in our dataset, represented by a binary
indicator (1 for polls at national level, 0 otherwise). This distinction is important because
not all polls are conducted nationally; some are state-specific and may reflect the
region's preference for a particular presidential candidate. Ignoring this distinction
may introduce significant bias into our model. By labeling each poll as national or
non-national, we can give higher weight to national polls, which are more representative
of overall public opinion, and thus improve the accuracy of our predictive model. @tbl-category shows that for the national_poll variable, non-national polls (coded as 0) account for 681 cases (75.0%), whereas national polls (coded as 1) represent 227 cases (25.0%). This distribution highlights a focus on likely voters and a predominance of non-national polling data in the sample.

### Population {#population}

The "population" variable in the dataset represents the status of each
entry. it is the abbreviated description of the respondent group, categorizing
participants as “lv” (likely voters) or “rv” (registered voters). They represent
different levels of participation and potential voting behavior. Likely voters
are individuals who are not only registered but are statistically more likely
to vote based on historical data or survey responses. In contrast, registered
voters include all individuals who are eligible to vote, regardless of their
likelihood of participation. @tbl-category shows that likely voters (lv) constitute the majority with 620 cases (68.3%), while registered voters (rv) make up 288 cases (31.7%).

@@

```{r}
#| label: fig-state
#| fig-cap: Count of polls by the US state 
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-pollster
#| fig-cap: Frequency of Polls by Pollster
#| echo: false
#| warning: false
#| message: false

@@

The linear regression models constructed for predicting support for
Kamala Harris and Donald Trump in the 2024 U.S. Presidential Election
utilized both unweighted and weighted configurations.The model summary of support for Harris is shown in @tbl-summary1, support for Trump is shown in @tbl-summary2. As shown in
@tbl-summary1 and @tbl-summary2, the intercepts for Harris and Trump,
estimated at 44.70% and 59.28% respectively in the unweighted models, while in weighted models, it's estimated at 45.2% and 54.8%,
represent their baseline support levels when all predictors are at
reference values. The high ( R\^2 ) values of 0.824 for Harris’s
unweighted model and 0.892 for her weighted model, along with similar
values of 0.849 and 0.916 for Trump, indicate that a significant portion
of the variation in support can be explained by the model's predictors.
The relatively low Root Mean Square Error (RMSE) values further suggest
that the models provide reasonably accurate predictions of support
levels.

@@

@fig-support-trend illustrate the predicted support for Kamala Harris
and Donald Trump from August 1, 2024, to November 2, 2024, using both
unweighted and weighted linear regression models. In the
@fig-support-trend-1 unweighted model, we see that Trump's predicted
support generally hovers between 40% and 50%, while Harris's support
ranges from 45% to 50%. Both candidates' support levels appear
relatively stable across the polling period, with minor fluctuations.
This model, which treats all polls equally, suggests a consistent lead for Harris, albeit within a fairly close range.

The weighted model @fig-support-trend-2 refines these predictions by
applying weights based on recency and sample size, which reduces the
influence of older or smaller polls. This approach results in smoother
trend lines for both candidates, with less apparent variability compared
to the unweighted model @fig-support-trend-1. Trump's support remains
around the 40% to 50% range, while Harris’s support stays in about 45% to 52%  range. The weighting emphasizes more recent and larger polls, likely
providing a more accurate reflection of current public opinion. Overall, both models consistently indicate a lead for Harris, with the weighted model slightly stabilizing the predictions. The average support for Harris and Trump are both consist for two models, which predicts that Harris will win on about 47%, leads Trump by about 1%.

@@

set.seed(233)
plan(multisession)
####### Prediction for Harris ######
new_data_harris <- data.frame(
  end_date = seq(
    min(harris_data$end_date),
    max(harris_data$end_date),
    length.out = 700
  ),
  state = factor(sample( c("National","Arizona", "Georgia", "Michigan", "Nevada", "North Carolina", "Pennsylvania", "Wisconsin"), size = 700, replace = TRUE)), 
  pollster = factor(sample(levels(harris_data$pollster), size = 700, replace = TRUE), 
                    levels = levels(harris_data$pollster)), 
  population = factor(sample(levels(harris_data$population), size = 700, replace = TRUE), 
                    levels = levels(harris_data$population)),
  sample_size = sample(1:2300, size = 700, replace = TRUE)
)|>
  mutate(
    national_poll = factor(if_else(state == "National", 1, 0)),
    recency = as.numeric(difftime(as.Date("2024-11-05"), end_date, units = "days")),
    sample_size_weight = pmin(sample_size / 2300, 1),
    recency_weight = exp(-recency * 0.1),
    combined_weight = recency_weight * sample_size_weight,
    candidate="Kamala Harris"
  )

####### Prediction for Trump ######
new_data_trump <- data.frame(
  end_date = seq(
    min(trump_data$end_date),
    max(trump_data$end_date),
    length.out = 700
  ),
  pollster = factor(sample(levels(trump_data$pollster), size = 700, replace = TRUE), 
                    levels = levels(trump_data$pollster)),
 population= factor(sample(levels(trump_data$population), size = 700, replace = TRUE), 
                    levels = levels(trump_data$population)),
   state = factor(sample( c("National","Arizona", "Georgia", "Michigan", "Nevada", "North Carolina", "Pennsylvania", "Wisconsin"), size = 700, replace = TRUE), 
                    levels = levels(trump_data$state)),  
  sample_size =sample(1:2300, size = 700, replace = TRUE)
)|>
  mutate(
   national_poll = factor(if_else(state == "National", 1, 0)),
      recency = as.numeric(difftime(as.Date("2024-11-05"), end_date, units = "days")),
    sample_size_weight = pmin(sample_size / 2300, 1),
    recency_weight = exp(-recency * 0.1),  
    combined_weight = recency_weight * sample_size_weight,
    candidate="Donald Trump"
  )
```

@@

# Filter combined_predictions to include only data from the recent date cutoff onward
combined_predictions<- bind_rows(new_data_harris, new_data_trump)
combined_predictions_filtered<-combined_predictions|>filter(end_date>=recent_date_cutoff)
# Determine the maximum end date for setting the x-axis limit
max_end_date <- as.Date("2024-11-2")
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| fig-cap: "Harris Kamala leads Donald Trump an average 2% in Predicted Support in the 2024 US Election (From August 1, 2024, to November 2, 2024)"
#| fig-subcap: ["Unweighted","Weighted"]
#| layout-ncol: 1
#| label: fig-support-trend
# Plotting with the filtered data
plan(multisession)
##unweighted
ggplot(combined_data_total,aes(x = end_date,y=pct, colour = candidate)) +
  geom_point(alpha = 0.1)+ 
  geom_smooth(data = combined_predictions_filtered,aes(x=end_date,y = predicted_unweighted), method = "lm",size = 1)+ 
  geom_line(data = combined_predictions_filtered,aes(x=end_date,y = predicted_unweighted),size = 0.2)+
  labs(
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate",caption = 
                   "Linear Regression Model: Candidate Support Percentage = national_poll + pollster + population + state") +
  scale_color_manual(values = c("Kamala Harris" = "blue", "Donald Trump" = "red")) +
  theme_classic() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 45, hjust = 1),strip.background = element_rect(colour = "black", fill = "white")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date)) +
  annotate("text", x = max(combined_predictions_filtered$end_date) -0.5, y = 48, 
           label = "47.5%", color = "blue", hjust = -0.2, size = 2) +  # Adjust size here
  annotate("text", x = max(combined_predictions_filtered$end_date) - 0.5, y = 46, 
           label = "46.5%", color = "red", hjust = -0.2, size = 2)

##weighted
ggplot(combined_data_total,aes(x = end_date,y=pct, colour = candidate,size = combined_weight)) +
  geom_point(alpha = 0.1)+ 
  geom_smooth(data = combined_predictions_filtered,aes(x=end_date,y = predicted_unweighted), method = "lm",size = 1)+ 
  geom_line(data = combined_predictions_filtered,aes(x=end_date,y = predicted_weighted), size = 0.2)+
  labs(x = "End Date of the Poll",
       y = "Predicted Support (%)",
       color = "Candidate",
       size="Weight",
       caption = "Weights: recency_weight * sample_size_weight \nLinear Regression Model: Candidate Support Percentage = national_poll + pollster + population + state") +
  scale_color_manual(values = c("Kamala Harris" = "blue", "Donald Trump" = "red"))  +
  theme_classic() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 45, hjust = 1),strip.background = element_rect(colour = "black", fill = "white")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date)) +
  annotate("text", x = max(combined_predictions_filtered$end_date) -0.5, y = 48, 
           label = "47.5%", color = "blue", hjust = -0.2, size = 2) +  # Adjust size here
  annotate("text", x = max(combined_predictions_filtered$end_date) - 0.5, y = 46, 
           label = "46.5%", color = "red", hjust = -0.2, size = 2)
```

@@

@fig-support-state compare the predicted support for Kamala Harris and
Donald Trump in key battleground states (Arizona, Georgia, Michigan,
Nevada, North Carolina, Pennsylvania, and Wisconsin) for the 2024 U.S.
election. @fig-support-state-1 uses an unweighted multiple linear
regression model, while @fig-support-state-2 applies a weighted model,
giving greater influence to polls with higher weights based on recency
and sample size. In both figures, Trump’s support, represented by red
lines, and Harris’s support, shown in blue, show close competition
across most states, with slight variations in trends across the polling
period from August to November 2024.

@@

The @fig-support-state-2 stabilizes these trends by emphasizing recent,
larger polls. Harris's support becomes more consistent, particularly in
Pennsylvania, Michigan, and Wisconsin, where she shows a clearer
lead. This weighted approach provides a more refined snapshot,
highlighting Harris’s likely advantage in these key states.

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| fig-subcap: ["Unweighted","Weighted"]
#| layout-ncol: 1
#| fig-cap: "Harris is likely to lead Trump in Pennsylvania, Michigan, and Wisconsin by Model Prediction"
#| label: fig-support-state
## unweighted
plan(multisession)
selected_states <- c("Arizona", "Georgia", "Michigan", "Nevada", "North Carolina", "Pennsylvania", "Wisconsin")
combined_data_total<-combined_data_total|>filter(state %in% selected_states)
combined_predictions_filtered<-combined_predictions_filtered|>filter(state %in% selected_states)
late<-combined_predictions_filtered[which.max(combined_predictions_filtered$end_date), ]
ggplot(combined_data_total,aes(x = end_date,y=pct, colour = candidate))+
  geom_point(alpha = 0.1)+ 
  geom_line(data = combined_predictions_filtered,aes(x=end_date,y = predicted_unweighted), size = 0.5)+
  facet_wrap(~ state,ncol = 4) +
  labs(
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate",
       caption = "Candidate Support Percentage = national_poll + pollster + population + state") +
  scale_color_manual(values = c("Kamala Harris" = "blue", "Donald Trump" = "red")) +
  theme_classic() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90, hjust = 1, size =6),
        legend.box = "vertical",
        strip.background = element_rect(colour = "black", fill = "white")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date))

## weighetd
ggplot(combined_data_total,aes(x = end_date,y=pct, colour = candidate,size = combined_weight)) +
  geom_point(alpha = 0.1)+ 
  geom_line(data = combined_predictions_filtered,aes(x=end_date,y = predicted_weighted), size = 0.5)+ # Smoothed trend line
  facet_wrap(~ state,ncol = 4) +
  labs(
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate",
       size="Weight",
       caption = "Weights: recency_weight * sample_size_weight\nCandidate Support Percentage = national_poll + pollster + population + state") +
  scale_color_manual(values = c("Kamala Harris" = "blue", "Donald Trump" = "red")) +
  theme_classic() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90, hjust = 1,size=6),
        legend.box = "vertical",
        strip.background = element_rect(colour = "black", fill = "white")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date)) 
```

@@

# Discussion {#sec-discussion}

@@

## Influence of Weighting Methodology in Analyzing Voters' Preference
